ActiveCA: Time Use Data from the General Social Survey of Canada to Study Active Travel

### 2024-12-13 paper/Manuscript-Data-Package.tex
Including labelled main files in the data paper

@@

\section{Methods}\label{methods}

For each selected cycle of the GSS surveys, we reviewed the episode
files to identify episodes of movement that involved walking or cycling.
This allowed us to select the activities immediately before and after
the movement episode. After that, we labeled the code variables with
their appropriate descriptions, identifying each origin and destination,
mode of travel, as well the province and urban classification of the
episode.

\section{How to use \{ActiveCA\}}\label{how-to-use-activeca}

@@

Considering GSS Cycles analyzed, we identified 21748 episodes that
recorded active travel episodes, with trip duration ranging from 0 to
900 minutes, to twelve different destinations. \texttt{ActiveCA}
includes all these episodes ready for analysis. Table \ref{tab:table-01}
presents descriptive statistics on walking and cycling trips between
1986 and 2015, including metrics such as the count of recorded trips
(count), and measures of trip duration in minutes: maximum (max), mean,
median, and minimum (min). The 1986 survey did not include bicycle
trips.

@@

\begin{longtable}[t]{>{}llcccccc}
\caption{\label{tab:table-01}\label{tab:table-01}Descriptive statistics for episodes with active transport records}\\
\toprule
\multicolumn{2}{c}{ } & \multicolumn{6}{c}{Year} \\
\cmidrule(l{3pt}r{3pt}){3-8}
Mode & Statistic & 1986 & 1992 & 1998 & 2005 & 2010 & 2015\\
\midrule
 & Count & 4347 & 1500 & 1670 & 5533 & 4379 & 3251\\
\nopagebreak
 & Maximum & 660 & 300 & 255 & 515 & 480 & 900\\
\nopagebreak
 & Mean & 21 & 19 & 11 & 12 & 12 & 17\\
\nopagebreak
 & Median & 10 & 10 & 5 & 10 & 8 & 10\\
\nopagebreak
 & Minimum & 1 & 1 & 1 & 0 & 0 & 5\\
\nopagebreak
\multirow[t]{-6}{*}{\raggedright\arraybackslash \textbf{Walking}} & Standard deviation & 31 & 25 & 17 & 16 & 17 & 27\\
\cmidrule{1-8}\pagebreak[0]
 & Count &  & 135 & 119 & 333 & 236 & 245\\
\nopagebreak
 & Maximum &  & 240 & 90 & 180 & 153 & 120\\
\nopagebreak
 & Mean &  & 31 & 21 & 19 & 21 & 24\\
\nopagebreak
 & Median &  & 20 & 15 & 15 & 15 & 15\\
\nopagebreak
 & Minimum &  & 5 & 2 & 1 & 1 & 5\\
\nopagebreak
\multirow[t]{-6}{*}{\raggedright\arraybackslash \textbf{Cycling}} & Standard deviation &  & 36 & 18 & 18 & 23 & 20\\
\bottomrule
\end{longtable}
\endgroup{}

Table \ref{tab:table-01} shows that the median values for walking trips
range between 5 and 10 minutes, while cycling trips have a consistent
median of 15 minutes since 1998. The table also highlights very high
maximum values, particularly for walking trips, with recorded episodes
exceeding 4 hours in all cases.

Table \ref{tab:table-02} and \ref{tab:table-03} provide descriptive
statistics for the two modes of transportation, split by destination
categories, from 1986 to 1998 and from 2005 to 2015, respectively. In
Table \ref{tab:table-02}, one can observed that in 1986 and 1992,
walking trips destined for \texttt{home} had the highest medians.
However, by 1998, the highest medians shifted to trips to
\texttt{work\ or\ school}, a transition that also occurred for cycling
trips between 1992 and 1998. Table \ref{tab:table-03} indicates that the
median duration for trips to \texttt{home} and \texttt{work\ or\ school}
remained at 10 minutes.

@@

\begin{ThreePartTable}
\begin{TableNotes}
\item \textit{Note: } 
\item * The symbols used in this table represent the following: 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(\%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination
\end{TableNotes}
\begin{longtable}[t]{ccccc>{}c|ccc>{}c|cccc}
\caption{\label{tab:table-02}\label{tab:table-02}Comparison of travel statistics by mode and destination: 1986, 1992, 1998}\\
\toprule
\multicolumn{2}{c}{ } & \multicolumn{4}{c}{1986} & \multicolumn{4}{c}{1992} & \multicolumn{4}{c}{1998} \\
\cmidrule(l{3pt}r{3pt}){3-6} \cmidrule(l{3pt}r{3pt}){7-10} \cmidrule(l{3pt}r{3pt}){11-14}
\multicolumn{1}{c}{\textbf{Destination}} & \multicolumn{1}{c}{\textbf{Mode*}} & \multicolumn{1}{c}{\textbf{Min*}} & \multicolumn{1}{c}{\textbf{Med*}} & \multicolumn{1}{c}{\textbf{Max*}} & \multicolumn{1}{c}{\textbf{(\%)*}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}}\\
\midrule
 & Home &  &  &  &  & 5 & 20 & 240 & 55.6 & 2 & 15.0 & 90 & 52.9\\
\nopagebreak
 & Other's home &  &  &  &  & 5 & 10 & 145 & 18.5 & 2 & 10.0 & 80 & 17.6\\
\nopagebreak
\multirow[t]{-3}{*}{\centering\arraybackslash Cycling} & Work or school &  &  &  &  & 5 & 15 & 45 & 25.9 & 5 & 20.0 & 75 & 29.4\\
\cmidrule{1-14}\pagebreak[0]
 & Home & 1 & 15 & 330 & 46.4 & 1 & 10 & 300 & 59.5 & 1 & 5.0 & 255 & 51.6\\
\nopagebreak
 & Other's home & 1 & 10 & 660 & 42.3 & 1 & 5 & 135 & 21.3 & 1 & 5.0 & 120 & 28.1\\
\nopagebreak
\multirow[t]{-3}{*}{\centering\arraybackslash Walking} & Work or school & 1 & 10 & 450 & 11.3 & 2 & 10 & 60 & 19.2 & 1 & 6.5 & 75 & 20.4\\
\bottomrule
\insertTableNotes
\end{longtable}
\end{ThreePartTable}
\endgroup{}

@@

@@

CAN WE PROVIDE PYTHON INTEGRATION? See here:

\url{https://github.com/paezha/idealista18/tree/master/Python}

@@

The value of \{ActiveCA\} lies in its transparency, accessibility, and
ease of use, which facilitates the addition of complementary data sets
in the future. \texttt{R} users can seamlessly explore GSS walking and
cycling episodes , with the option to suggest enhancements to the
package as needed. This article adopts the structure proposed by
Anastasia and Páez \citeyearpar{soukhov2023}, whose work provided
essential guidance for the creation of this package. Similarly, we aim
to contribute to the academic community by promoting transparent
research practices that encourage replication and innovation in related
fields. We believe that \{ActiveCA\} will serve as a basis for further
research on GSS and for the integration of additional data by the
authors or the wider open source community.

>>>>

Table \ref{tab:main-2015-unprocessed} shows the first ten rows and first
six variables of the GSS PUMF 2015 main file (Cycle 29). Each row in the
table correspond to a survey respondent, while the columns refer the
following information: record identification (\texttt{PUMFID}), the
person's weight (\texttt{WGHT\_PER}), the month the survey data was
collected (\texttt{SURVMNTH}), the respondent's age group
(\texttt{AGEGR10}), the respondent's sex (\texttt{SEX}), and the
respondent's marital status (\texttt{MARSTAT}).

The main file of the 2015 GSS surveys includes a total of 17,390
respondents, representing 29,766,399 individuals and 848 variables. For
discrete variables, Statistics Canada has assigned specific codes to the
possible values, with each code accompanied by a label. For instance, in
the case of the variable \texttt{SURVMNTH}, a value of \texttt{1} means
\texttt{January\ 2016}, 2 means \texttt{February\ 2016}, 3 corresponds
to \texttt{March\ 2016}, and so on.

As shown in Table \ref{tab:main-2015-unprocessed}, the variables are not
labeled. Adittionaly, the format of the tables (comma-separated values)
does not allow for the specification of variable types (whether a
variable is continuous or discrete), which can lead to mistakes analysts
who have limited experience working with PUMFs.

\begingroup\fontsize{8}{10}\selectfont

\begin{ThreePartTable}
\begin{TableNotes}
\item \textit{Note: } 
\item Legend: `PUMFID`: record identification. `WGHT\_PER`:  person weight. `SURVMNTH`: survey month of data collection. `AGEGR10`: age group of the respondent. `SEX`: sex of the respondent. `MARSTAT`: marital status of the respondent.
\end{TableNotes}
\begin{longtable}[t]{cccccc}
\caption{\label{tab:gss-main-file-2015}\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the main file of the 2015 GSS.}\\
\toprule
PUMFID & WGHT\_PER & SURVMNTH & AGEGR10 & SEX & MARSTAT\\
\midrule
10000 & 616.6740 & 7 & 5 & 1 & 5\\
10001 & 8516.6140 & 7 & 5 & 1 & 1\\
10002 & 371.7520 & 1 & 4 & 2 & 1\\
10003 & 1019.3135 & 3 & 6 & 2 & 5\\
10004 & 1916.0708 & 9 & 2 & 1 & 6\\
\addlinespace
10005 & 1952.2015 & 4 & 1 & 1 & 6\\
10006 & 5761.5528 & 8 & 1 & 1 & 6\\
10007 & 466.0426 & 6 & 5 & 2 & 3\\
10008 & 2479.2991 & 2 & 2 & 2 & 1\\
10009 & 1436.1641 & 8 & 6 & 1 & 3\\
\bottomrule
\insertTableNotes
\end{longtable}
\end{ThreePartTable}
\endgroup{}

@@

Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an
example of the main file structure, Table \ref{tab:ep-2015-unprocessed}
presents the first seven episodes for the record identification number
\texttt{10041} and some variables from the GSS PUMF 2015 episode file
(Cycle 29). Each row in the table corresponds to an episode associated
with the specified record identification (\texttt{PUMFID\ =\ 10041}),
including the episode's weight (\texttt{WGHT\_EPI}), episode number
(\texttt{EPINO}), activity code (\texttt{TUI\_01}), episode duration
(\texttt{DURATION}), and episode location (\texttt{LOCATION}).

In total, the episode file of the 2015 GSS surveys contains 274,108
records, representing 461,837,622 episodes and 527 variables detailing
the episodes. Similar to the main file, Statistics Canada has created
codes for the discrete variables, with each value corresponding to a
label.

In the case illustrated in Table \ref{tab:ep-2015-unprocessed}, this
respondent began the diary description by sleeping at home
(\texttt{TUI\_01\ =\ 1} and \texttt{LOCATION\ =\ 300}) for 210 minutes,
followed by 40 minutes of personal hygiene (\texttt{TUI\_01\ =\ 2}). The
respondent then spent 15 minutes on personal care activities, such as
getting ready for school, supervising homework, reading, playing,
reprimanding, or providing educational or emotional support, as
indicated by \texttt{TUI\_01\ =\ 27}. Next, they recorded a travel
episode, walking for 15 minutes (\texttt{TUI\_01\ =\ 7} and
\texttt{LOCATION\ =\ 315}), where both the origin and destination were
their home. Such trips, where the journey starts and finishes at home,
can be classified as recreational or leisure trips. Next, the respondent
spent 3 hours searching for a job (\texttt{TUI\_01\ =\ 9}), took a
15-minute lunch break (\texttt{TUI\_01\ =\ 6}), and then cleaned the
house (\texttt{TUI\_01\ =\ 18}) for two hours. Table
\ref{tab:ep-2015-unprocessed} displays only six variables out of the 527
available. As shown, since the dataset does not label the variable
values, decoding them can be both time-consuming and challenging.

\begingroup\fontsize{8}{10}\selectfont

\begin{ThreePartTable}
\begin{TableNotes}
\item \textit{Note: } 
\item Legend: `PUMFID`: record identification. `EPINO`: episode number. WGHT\_EPI: episode's weight. TUI\_01: activity code. DURATION: episode's duration. LOCATION: episode's location.
\end{TableNotes}
\begin{longtable}[t]{cccccc}
\caption{\label{tab:gss-epi-file-2015}\label{tab:ep-2015-unprocessed}Visualization of the seven first episodes of the record number `10041`.}\\
\toprule
PUMFID & EPINO & WGHT\_EPI & TUI\_01 & DURATION & LOCATION\\
\midrule
10041 & 1 & 1353.818 & 1 & 210 & 300\\
10041 & 2 & 1353.818 & 2 & 40 & 300\\
10041 & 3 & 1353.818 & 27 & 15 & 300\\
10041 & 4 & 1353.818 & 7 & 15 & 315\\
10041 & 5 & 1353.818 & 9 & 180 & 300\\
\addlinespace
10041 & 6 & 1353.818 & 6 & 15 & 300\\
10041 & 7 & 1353.818 & 18 & 120 & 300\\
\bottomrule
\insertTableNotes
\end{longtable}
\end{ThreePartTable}
\endgroup{}

\section{Data extraction}\label{data-extraction}

For each selected cycle of the GSS surveys, we reviewed the episode
files to identify episodes of movement that involved walking or cycling.
This allowed us to select the activities immediately before and after
the movement episode. After that, we labeled the code variables with
their appropriate descriptions, identifying each origin and destination,
mode of travel, time spent in the active trip, as well the province and
urban classification of the episode.

The active trips (walking and cycling) were identified by their
corresponding activity codes, accounting for differences between the
survey cycles. Identifying the preceding and succeeding activities of
each activity episode helps to understand the motivation for the trip
and also determine the origin and destination of the trips.

For each GSS main file, we selected socioeconomic variables to help
profile the individuals engaged in active trip episodes. We included key
socioeconomic variables such as the respondent's age group, sex, marital
status, and number of children, among others. After this, as we did with
the episode files, we labeled the code variables with their appropriate
descriptions.

\section{\texorpdfstring{How to use
\texttt{ActiveCA}}{How to use ActiveCA}}\label{how-to-use-activeca}

This section presents some potential applications of the
\texttt{ActiveCA} R package. In fact, we expect that the application of
this package to extend beyond our pre-imagined range of uses.

To start the demonstration, it is first necessary to install the
package. The code chunk below demonstrates how to install the
\texttt{ActiveCA} package in R:

\begin{verbatim}
if(!require("ActiveCA", character.only = TRUE)) {
      remotes::install_github("dias-bruno/ActiveCA")
}
\end{verbatim}

After installing the \texttt{ActiveCA} R package, you can load it using
the following code:

\begin{verbatim}
library(ActiveCA)
\end{verbatim}

By running \texttt{?ActiveCA} in \texttt{R}, you can access the complete
package documentation. The documentation includes a description of the
package, information about the authors and maintainers, and an overview
of all pre-processed data sets included in the package.

Figure \ref{fig:walking_documentation} displays the documentation for
the walking episodes from the 2015 GSS (\texttt{walking\_2015.rda}). The
documentation provides a brief description of the dataset, instructions
on how to load the package, as well as details about the file format and
descriptions of all variables (already labeled).

{\centering \includegraphics[width=1\linewidth]{Manuscript-figures/walking-2015-documentation}

\caption{Documentation for the dataset of walking episodes from 2015.}\label{fig:walking_documentation}
\end{figure}

The chunk below shows how to load the walking episodes from the 2015 GSS
(Cycle 29):

\begin{verbatim}
data(walking_2015)
\end{verbatim}

The GSS surveys apply a probability sampling methodology, in which each
episode or person selected in the sample represents several other
episodes or persons not in the sample. The number of episodes and
persons represented by a episode or person is determined by the weight
or weighting factor. Because of this, every estimates of the number of
episodes or persons need to be calculated applying the corresponding
weighting factors.

For instance, if someone wants to calculate the percentage of
respondents from the 2015 GSS survey with active travel episodes, it is
necessary to account for the person's weights. In 2015, the weight
variable is represented by \texttt{WGHT\_PER.} The code chunk below
demonstrates how to obtain the percentage of people with active travel
episodes by age group. It uses the dplyr package to manipulate the data.

The process begins by creating a dataset that sums the population by age
group. Then, it joins the 2015 episodes with the 2015 main file. Note
that in both operations, the code sums the person weight variable
(\texttt{WGHT\_PER}) to obtain the correct population values. After
this, a new dataset, \texttt{Active\_percentage}, is created by merging
both previous datasets. The percentage of active travel episodes by age
group is calculated by dividing the total population by the population
with active trip episodes, then multiplying by 100.

The result shows that the age group with the highest share of active
trips is those between 15 and 20 years old, with almost 37\%, followed
by those between 25 and 34 years old with around 33\%. There is a
significant drop in percentage for the following groups, with the
percentage falling to between 15\% and 18\% for the remaining age
groups.

\begin{verbatim}
# Load the dplyr library for data manipulation
library('dplyr')

# Load the episodes dataset
data(gss_episodes)

# Calculate the total population for each age group
Total_population <- gss_main_2015 %>% 
  group_by(AGEGR10) %>% 
  summarise(Total_population  = sum(WGHT_PER))

# Calculate the active population for each age group
Active_population <- gss_episodes %>% 
      filter(YEAR == 2015) %>% 
  left_join(gss_main_2015, by = c("PUMFID"="PUMFID")) %>% 
  group_by(AGEGR10) %>% 
  summarise(Active_population  = sum(WGHT_PER))

# Combine total and active population and calculate the percentage
Active_percentage <- Total_population %>% 
  left_join(Active_population, by = c("AGEGR10" = "AGEGR10")) %>%
  mutate(Percentage = round(Active_population/Total_population  * 100, 2))

# Display the final result
Active_percentage
\end{verbatim}

\begin{verbatim}
# A tibble: 7 x 4
  AGEGR10           Total_population Active_population Percentage
  <ord>                        <dbl>             <dbl>      <dbl>
1 15 to 24 years            4511131.          1641468.       36.4
2 25 to 34 years            4956386.          1644033.       33.2
3 35 to 44 years            4734506.           847770.       17.9
4 45 to 54 years            5136125.           790667.       15.4
5 55 to 64 years            4831306.           767064.       15.9
6 65 to 74 years            3283969.           559025.       17.0
7 75 years and over         2312976.           384359.       16.6
\end{verbatim}

\subsection{Processed data set}\label{processed-data-set}

Table \ref{tab:processed-obs} displays the total number of observations
processed for main and episode files. For the main files, a total of
89,331 observations were processed, referring to all observations from
the Time Use Surveys from 1986 to 2015, that together represents more of
149,389,839 respondents (29,766,399 for 2015, 28,075,610 for 2010,
26,095,819 for 2005, 24,260,137 for 1998, 21,294,313 for 1992, and for
2005, 24,260,137 for 1998, and 21,294,313 for 1986). Table X also
presents the total cases of active trips episodes identified. In total
21,748 observations with register of active travel activity. Together,
these observations account for NA episodes (0 for 2015, 0 for 2010, 0
for 2005, 0 for 1998, 0 for 1992, and for 2005, 0 for 1998, and 0 for
1986).

\begin{longtable}[t]{lrrr}
\caption{\label{tab:table_df_processed}\label{tab:processed-obs}Total number and weighted sum of observations processed.}\\
\toprule
\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Observations} \\
\cmidrule(l{3pt}r{3pt}){3-4}
Survey & Year & Count & Weighted\\
\midrule
 & 2015 & 17390 & 29766399\\
\nopagebreak
 & 2010 & 15390 & 28075610\\
\nopagebreak
 & 2005 & 19597 & 26095819\\
\nopagebreak
 & 1998 & 10749 & 24260137\\
\nopagebreak
 & 1992 & 9815 & 21294313\\
\nopagebreak
\multirow[t]{-6}{*}{\raggedright\arraybackslash Main} & 1986 & 16390 & 19897562\\
\cmidrule{1-4}\pagebreak[0]
 & 2015 & 3496 & 6634387\\
\nopagebreak
 & 2010 & 4615 & 8516753\\
\nopagebreak
 & 2005 & 5866 & 7583838\\
\nopagebreak
 & 1998 & 1789 & 3606987\\
\nopagebreak
 & 1992 & 1635 & \\
\nopagebreak
\multirow[t]{-6}{*}{\raggedright\arraybackslash Episode} & 1986 & 4347 & 8241196\\
\bottomrule
\end{longtable}
\endgroup{}

Table \ref{tab:main-2015-processed} shows the first ten rows and first
six variables of the GSS PUMF 2015 main file (Cycle 29), displayed in
\ref{tab:main-2015-unprocessed} before our processing. Table
\ref{tab:ep-2015-processed} presents the walking episodes for the record
identification number \texttt{10041} from the GSS PUMF 2015 episode file
(Cycle 29), previously displayed in Table \ref{tab:ep-2015-unprocessed}.
Only active travel episodes appear in Table \ref{tab:ep-2015-processed}
since the observations were filtered to select cases with walking or
cycling episodes. For both cases, Tables \ref{tab:main-2015-processed}
and \ref{tab:ep-2015-processed} contain labeled variables, facilitating
the interpretation of the data.

\begingroup\fontsize{8}{10}\selectfont

\begin{ThreePartTable}
\begin{TableNotes}
\item \textit{Note: } 
\item Legend: `PUMFID`: record identification. `WGHT\_PER`:  person weight. `SURVMNTH`: survey month of data collection. `AGEGR10`: age group of the respondent. `SEX`: sex of the respondent. `MARSTAT`: marital status of the respondent.
\end{TableNotes}
\begin{longtable}[t]{cccccc}
\caption{\label{tab:gss-processed-file-2015}\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the main file of the 2015 GSS.}\\
\toprule
PUMFID & WGHT\_PER & SURVMNTH & AGEGR10 & SEX & MARSTAT\\
\midrule
10000 & 616.6740 & July & 55 to 64 years & Male & Divorced\\
10001 & 8516.6140 & July & 55 to 64 years & Male & Married\\
10002 & 371.7520 & January & 45 to 54 years & Female & Married\\
10003 & 1019.3135 & March & 65 to 74 years & Female & Divorced\\
10004 & 1916.0708 & September & 25 to 34 years & Male & Single, never married\\
\addlinespace
10005 & 1952.2015 & April & 15 to 24 years & Male & Single, never married\\
10006 & 5761.5528 & August & 15 to 24 years & Male & Single, never married\\
10007 & 466.0426 & June & 55 to 64 years & Female & Widowed\\
10008 & 2479.2991 & February & 25 to 34 years & Female & Married\\
10009 & 1436.1641 & August & 65 to 74 years & Male & Widowed\\
\bottomrule
\insertTableNotes
\end{longtable}
\end{ThreePartTable}
\endgroup{}

\begingroup\fontsize{8}{10}\selectfont

\begin{ThreePartTable}
\begin{TableNotes}
\item \textit{Note: } 
\item Legend: `PUMFID`: record identification. `EPINO`: episode number. WGHT\_EPI: episode's weight. TUI\_01: activity code. DURATION: episode's duration. LOCATION: episode's location.
\end{TableNotes}
\begin{longtable}[t]{rrlrlll}
\caption{\label{tab:gss-processed-file-2015}\label{tab:ep-2015-processed}Visualization of active travel episodes of the record number `10041` for the 2015 GSS survey.}\\
\toprule
PUMFID & WGHT\_EPI & Activity & Duration & Origin & Destination & Mode\\
\midrule
10041 & 1353.818 & Transport to or from activity & 15 & Home & Home & Walking\\
\bottomrule
\insertTableNotes
\end{longtable}
\end{ThreePartTable}
\endgroup{}

@@

Considering GSS Cycles analyzed, we identified 21,748 episodes that
recorded active travel episodes, with trip duration ranging from 0 to
900 minutes, to twelve different destinations. \texttt{ActiveCA}
includes all these episodes ready for analysis. Table \ref{tab:table-01}
presents descriptive statistics on walking and cycling trips between
1986 and 2015, with measures of trip duration in minutes: maximum (max),
mean, median, and minimum (min). The 1986 survey did not include bicycle
trips.

@@

\begin{longtable}[t]{>{}llcccccc}
\caption{\label{tab:table-01}\label{tab:table-01}Descriptive statistics for episodes with active transport records}\\
\toprule
\multicolumn{2}{c}{ } & \multicolumn{6}{c}{Year} \\
\cmidrule(l{3pt}r{3pt}){3-8}
Mode & Statistic & 1986 & 1992 & 1998 & 2005 & 2010 & 2015\\
\midrule
 & Maximum & 660 & 300 & 255 & 515 & 480 & 900\\
\nopagebreak
 & Mean & 21 &  & 12 & 12 & 13 & 18\\
\nopagebreak
 & Median & 15 & 10 & 5 & 10 & 10 & 10\\
\nopagebreak
 & Minimum & 1 & 1 & 1 & 0 & 0 & 5\\
\nopagebreak
\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Walking}} & Standard deviation & 31 & 25 & 17 & 16 & 17 & 27\\
\cmidrule{1-8}\pagebreak[0]
 & Maximum &  & 240 & 90 & 180 & 153 & 120\\
\nopagebreak
 & Mean &  & 28 & 24 & 20 & 19 & 25\\
\nopagebreak
 & Median &  & 15 & 15 & 15 & 10 & 20\\
\nopagebreak
 & Minimum &  & 5 & 2 & 1 & 1 & 5\\
\nopagebreak
\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Cycling}} & Standard deviation &  & 36 & 18 & 18 & 23 & 20\\
\bottomrule
\end{longtable}
\endgroup{}

Table \ref{tab:table-01} shows that, in general, the median values for
walking trips is 10 minutes, except for 1998 when the median was 5
minutes. In the case of cycling trips, the duration fluctuated over the
years, ranging from 10 to 20 minutes. The table also highlights very
high maximum values, particularly for walking trips, with recorded
episodes exceeding 4 hours in all cases.

Table \ref{tab:table-02} and \ref{tab:table-03} provide descriptive
statistics for the two modes of transportation, split by destination
categories, from 1986 to 1998 and from 2005 to 2015, respectively. In
Table \ref{tab:table-02}, one can observed that in 1986 and 1992,
walking trips destined for \texttt{home} had the highest medians.
However, by 1998, the highest medians shifted to trips to
\texttt{work\ or\ school}, a transition that also occurred for cycling
trips between 1992 and 1998. Table \ref{tab:table-03} indicates that the
median duration for walking trips to \texttt{home} and
\texttt{work\ or\ school} remained at 10 minutes.

@@

\begin{ThreePartTable}
\begin{TableNotes}
\item \textit{Note: } 
\item 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(\%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination
\end{TableNotes}
\begin{longtable}[t]{ccccc>{}c|ccc>{}c|cccc}
\caption{\label{tab:table-02}\label{tab:table-02}Comparison of travel statistics by mode and destination: 1986, 1992, 1998}\\
\toprule
\multicolumn{2}{c}{ } & \multicolumn{4}{c}{1986} & \multicolumn{4}{c}{1992} & \multicolumn{4}{c}{1998} \\
\cmidrule(l{3pt}r{3pt}){3-6} \cmidrule(l{3pt}r{3pt}){7-10} \cmidrule(l{3pt}r{3pt}){11-14}
\multicolumn{1}{c}{\textbf{Destination}} & \multicolumn{1}{c}{\textbf{Mode}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}}\\
\midrule
 & Home &  &  &  &  & 5 & 20 & 240 & 55.7 & 2 & 15 & 90 & 51.6\\
\nopagebreak
 & Other's home &  &  &  &  & 5 & 10 & 145 & 19.7 & 2 & 10 & 80 & 15.7\\
\nopagebreak
\multirow[t]{-3}{*}{\centering\arraybackslash Cycling} & Work or school &  &  &  &  & 5 & 15 & 45 & 24.7 & 5 & 25 & 75 & 32.8\\
\cmidrule{1-14}\pagebreak[0]
 & Home & 1 & 15 & 330 & 46.3 & 1 & 15 & 300 &  & 1 & 5 & 255 & 51.9\\
\nopagebreak
 & Other's home & 1 & 10 & 660 & 42.6 & 1 & 5 & 135 &  & 1 & 5 & 120 & 27.8\\
\nopagebreak
\multirow[t]{-3}{*}{\centering\arraybackslash Walking} & Work or school & 1 & 10 & 450 & 11.1 & 2 & 10 & 60 &  & 1 & 10 & 75 & 20.3\\
\bottomrule
\insertTableNotes
\end{longtable}
\end{ThreePartTable}
\endgroup{}

@@

\texttt{ActiveCA} also enables obtaining insights from the main
processed files. Figure \ref{fig:figure-stress} present how the level of
stress varied among respondents depending on their marital status in
2015. According to this plot, married respondents reported the highest
level of stress, with 17\% of possible cases.

{\centering \includegraphics[width=1\linewidth]{Manuscript-figures/main_stress_figure}

\caption{Level of stress among respondents of different marital statuses (2015).}\label{fig:figure-stress}
\end{figure}

@@

@@

The value of \{ActiveCA\} lies in its transparency, accessibility, and
ease of use, which facilitates the addition of complementary data sets
in the future. \texttt{R} users can seamlessly explore GSS walking and
cycling episodes, with the option to suggest enhancements to the package
as needed. This article adopts the structure proposed by Anastasia and
Páez \citeyearpar{soukhov2023}, whose work provided essential guidance
for the creation of this package. Similarly, we aim to contribute to the
academic community by promoting transparent research practices that
encourage replication and innovation in related fields. We believe that
\{ActiveCA\} will serve as a basis for further research on GSS and for
the integration of additional data by the authors or the wider open
source community.
